---
title: "p8105_hw3_as5457"
author: "Sunny Siddique"
date: "October 8, 2018"
output: github_document
---

Setting global settings for the format of the graphs generated for this homework. 
```{r}

#Loading the tidyverse package. 
library(tidyverse)

#Ensuring that the generated graphs are of reasonable height and width.
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

#Ensuring that the background of the graphs are bw and legend is printed at the bottom. 
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

------------------------------------------------------------------------------------------------

#Problem 1

**Load the BRFSS data**

First, do some data cleaning:

- Format the data to use appropriate variable names;
- Focus on the “Overall Health” topic
- Include only responses from “Excellent” to “Poor”
- Organize responses as a factor taking levels from “Excellent” to “Poor”

```{r}
#Loading the dataset from the p8105.datasets package
library(p8105.datasets)
data("brfss_smart2010")

#Defining a new dataset called brfss_smart from the original dataset
brfss_smart = brfss_smart2010 %>% 
  
#Cleaning variable names
janitor::clean_names(dat = .) %>% 
  
#Keeping only rows where the topic was "Overall Health"
  filter(topic == "Overall Health")
  
#Organizing responses as a factor taking levels from "Excellent" to "Poor"

brfss_smart$response = as.factor(brfss_smart$response)
class(brfss_smart$response)
brfss_smart %>%
  mutate(response = forcats::fct_reorder(response, display_order))
class(brfss_smart$response)

```


**Using this dataset, do or answer the following (commenting on the results of each):**

- In 2002, which states were observed at 7 locations?

The states that were observed at 7 locations were CT, FL, NC. 

```{r}
brfss_smart %>%
  filter(year == 2002) %>% 
  separate(locationdesc, into = c("state", "county"), sep = " - ") %>% 
  distinct(state, county) %>% 
  count(state) %>% 
  filter (n == 7)
```


- Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r}
spaghetti = brfss_smart %>% 
  group_by (year, locationabbr) %>% 
  summarize (n = n()) %>% 
  ggplot(aes(x = year, y = n, color = locationabbr)) +
  geom_line() + 
  labs (
    title = "Number of Observations in Each State by Year",
        x = "year",
        y = "number of observations"
  )
spaghetti
```


- Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.


```{r}

year02 = brfss_smart %>% 
  filter (year == 2002)
year06 = brfss_smart %>%
  filter (year == 2006)
year10 = brfss_smart %>% 
  filter (year == 2010)
filteredyear = bind_rows(year02, year06, year10)

filteredyear %>% 
  filter(locationabbr == "NY") %>% 
  filter(response == "Excellent") %>% 
  group_by(year) %>% 
  summarize(mean_prop = mean(data_value),
            std_prop = sd(data_value))
```

- For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

First converting year into a factor variable

```{r}
brfss_smart$year=as.factor(brfss_smart$year)
```


```{r}
brfss_smart %>% 
  group_by (response, locationabbr, year) %>% 
  summarize(avg_prop = mean (data_value)) %>% 
  ggplot(aes(x = year, y = avg_prop)) + 
  geom_boxplot(alpha = .5) +
  facet_grid(. ~ response) +
  labs (
    title = "Distribution of state-level averages over time",
        x = "year",
        y = "average proportion"
  )
```



# Problem 2
Importing the instacart data

```{r}
#Loading the dataset from the p8105.datasets package
library(p8105.datasets)
data("instacart")

#Defining a new dataset called brfss_smart from the original dataset
insta = instacart %>% 
  
#Cleaning variable names
janitor::clean_names(dat = .)
```



The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.



Then, do or answer the following (commenting on the results of each):

** How many aisles are there, and which aisles are the most items ordered from? **
There are 134 aisles.

```{r}
unique_aisle = unique(insta$aisle_id)
length(unique_aisle)
```

```{r}
items_ordered = insta %>%
  group_by(aisle) %>% 
  summarize(item_count = n()) %>% 
  mutate(item_count = desc(item_count))

items_sorted = items_ordered %>% 
  sort(item_count, decreas)
  
  

  mutate(order(item_count, decreasing = TRUE))
sort(items_ordered$item_count, decreasing = FALSE)
```


Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).






